\fontfamily{phv}\selectfont{Causal Discovery methods for Effective Connectivity}

### 2013-09-27 docs/proposal/doc.tex
Added small section on derivation of structural and functional networks.

\begin{abstract}
Lorem Ipsum Dolor Sit Amet.
\keywords{Keywords}
\end{abstract}

@@

The field of brain connectivity finds its roots in the early 1990s \cite{friston1993functional, friston1994}, though because of more recent developments in the field of Artificial Intelligence and Machine Learning it is possible to do more thorough analysis \cite{vandenheuvel2010}.
One type of connectivity analysis involves finding how brain region X causes what region Y does. This is called \emph{effective connectivity}.
More concretely: how does the measured fMRI signal, representing an underlying neuronal process, determine what other neuronal processes are activated.
Causal analysis is also relatively new and its application in brain connectivity analysis still faces many challenges that require further research \cite{ramsey2010}.

@@

@@

Another causal discovery method is the BCCD-algorithm \cite{claassen2012}, which gives similar results as PC, and is designed in an attempt to improve PC.
This algorithm has some useful properties; it is a Bayesian approach and as such, it is more robust against noise.
The authors have shown the method to perform as good as or better than other causal discovery algorithms such as PC on representative datasets.
As is inherent to Bayesian methods, or approximations of Bayesian methods, the resulting network graph is not a  single structure but rather a distribution over possible structures.
Concretely, this provides an uncertainty measure over edges in the network, giving additional information compared to PC.

\subsection*{Data acquisition \& analysis}
We will apply the mentioned methods on structural and functional data gathered from human subjects.
We use the same functional and structural data as in Hinne et al. \cite{hinne2013}, of which six subjects were selected at random.
To find effective connectivity, both data sources are needed.
Structural connectivity of these subjects can be derived from DTI-measurements, using probabilistic tractography \cite{hinne2013}.
Functional connectivity can be extracted from fMRI-measurement, possibly in combination with the structural 
connectivity.
Functional connectivity can be extracted from fMRI-measurement, possibly in combination with the structural connectivity.

>>>>

\begin{abstract}
We propose a new application for causal discovery methods on human brain networks in order to derive effective connectivity.
The directive nature of effective brain networks suggest causal methods can be used.
As current methods for effective connectivity suffer from computational difficulties and often cannot cope with latent variables, we propose an application of PC-algorithm and the Bayesian BCCD-algorithm \cite{claassen2012} to estimate effective connectivity.
Both methods will be applied on six healthy subjects of which fMRI data and structural networks are readily available.
\keywords{Effective connectivity, Causal Discovery}
\end{abstract}

@@

The field of brain connectivity finds its roots in the early 1990s \cite{friston1993functional, friston1994}, though because of more recent developments in the field of Artificial Intelligence and Machine Learning it is possible to do more thorough analysis \cite{vandenheuvel2010}.
One type of connectivity analysis involves finding how brain region X causes what region Y does. This is called \emph{effective connectivity}.
More concretely: how does the measured fMRI signal, representing an underlying neuronal process, determine what other neuronal processes are activated.
Research regarding effective is relatively new and its application still faces many challenges that require further research \cite{ramsey2010}.
The directed character in effective connectivity seems similar to traditional causal relations.
In what follows, we suggest to use causal discovery methods for deriving effective connectivity.

@@

Hinne et al. \cite{hinne2013} proposed a Bayesian method for estimating structural networks based on Diffusion Weighted MRI (DWI) and probabilistic Tractography.
The location of white-matter tracts does not become immediately clear from imagery alone.
Other methods rely on thresholds to estimate location of said tracts.
The proposed method results in a measure of uncertainty about where a hypothesised connection will terminate and so provides a clearly interpretable network structure as result.

@@

Another causal discovery method is the Baysian Constraint-based Causal Discovery (BCCD) algorithm \cite{claassen2012}, which gives similar results as PC, and is designed in an attempt to improve PC.
This algorithm has some useful properties; it is a Bayesian approach and as such, it is more robust against noise.
The authors have shown the method to perform as good as or better than other causal discovery algorithms such as PC on representative datasets.
As is inherent to Bayesian methods, or approximations of Bayesian methods, the resulting network graph is not a  single structure but rather a distribution over possible structures.
Concretely, this provides an uncertainty measure over edges in the network, giving additional information compared to PC.

\subsection*{Data acquisition \& analysis}
We will apply the mentioned methods on structural and functional data gathered from human subjects.
For this we are largely dependent on data collected by others.
Hinne et al. has provided fMRI time-series data and structural network data as used and derived in \cite{hinne2013}, of which six subjects were selected at random.
We can estimate functional connectivity from fMRI-measurement in combination with structural 
networks according to a procedure currently submitted by Hinne et al. \cite{hinne2013submitted}.
To find effective connectivity, both structural networks and functional data is needed.
